Geometric Filters for Metric Spaces

### 2018-10-17 paper/metric_space_exploration.tex
Copy proof from Bose et al for approximate greedy spanner

\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\renewcommand{\S}{\mathbb{S}}
%\newcommand{\Wlog}{W.l.o.g.\ }
%\newcommand{\wlog}{w.l.o.g.\ }
\newcommand{\eps}{\epsilon}
\newcommand{\orient}{O}
\newcommand{\cpro}{\Pi}
\newcommand{\fib}{\mathrm{Fib}}

\newcommand{\parent}{\mathrm{par}}
\newcommand{\point}{p}
\newcommand{\level}{\ell}
\newcommand{\diameter}{\mathrm{diam}}
\newcommand{\pointset}{P}
\newcommand{\metricspace}{\mathcal{M}}
\newcommand{\metric}{\delta}
\newcommand{\approxmetric}{\gamma}
\newcommand{\complexity}{C_{\metric}}
\newcommand{\doublingdimension}{\Delta}

@@

\subsection{WSPD}

@@

>>>>

\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\dappr}{\tilde{d}}
%\newcommand{\Wlog}{W.l.o.g.\ }
%\newcommand{\wlog}{w.l.o.g.\ }
\newcommand{\eps}{\epsilon}
\newcommand{\orient}{O}
\newcommand{\cpro}{\Pi}
\newcommand{\fib}{\mathrm{Fib}}

\newcommand{\parent}{\mathrm{par}}
\newcommand{\point}{p}
\newcommand{\level}{\ell}
\newcommand{\diam}{\mathrm{diam}}
\newcommand{\pointset}{P}
\newcommand{\metricspace}{\mathcal{M}}
\newcommand{\metric}{\delta}
\newcommand{\approxmetric}{\gamma}
\newcommand{\complexity}{C_{\metric}}
\newcommand{\doublingdimension}{\Delta}

@@

\subsection{WSPD}
Given $s > 1$, two disjoint subsets $A, B$ of a metric space $(X, d)$ are called $s$-\textit{well-separated},
if 
\[
\forall a \in A \, \forall b \in B \, d(a, b) \geq s \max(\diam(A), \diam(B))
\]
A well-separated pair decomposition (WSPD) is a set of unordered pairs of sets $\{ \{A_1, B_1 \}, 
\dots, \{A_s, B_s\} \}$ such that each pair $\{A_i, B_i\}$ is $s$-well-separated, and for every unordered pair $\{a, b\}$ of distinct points of $X$ there exists a unique $j$ such that $a \in A_j$ and $b \in B_j$.
The notion of WSPD was introduced by Callahan and Kosaraju \cite{cal-kos-wspd}.

@@

\subsection{Approximate Greedy Spannner}

Suppose that we have fast algorithms for computing $\lambda$-approximation of $d(x,y)$.
In other words, for every pair of points $x,y$ we can compute an interval $[c, \lambda c]$ that is guaranteed to contain 
the true distance $d(x,y)$.  Note that for every interval of this form there exists 
a unique $k \in \{1, \dots, N-1 \}$ such that $[c, \lambda c] \subset [\lambda^k, \lambda^{k+2}]$. 
Indeed, $k = \lceil \log_{\lambda}(c) \rceil$.
We also assume that we know the range of all possible values of $k$ and they are all non-negative,
$k = 1\dots N$. We call the intervals $[\lambda^k, \lambda^{k+2}]$ \textit{buckets}, even though they are overlapping.
Now we can formulate the algorithm for finding a spanner.

\begin{algorithmic}
\label{alg:local_init}
\Function{Approximate greedy spanner}{$rho$}
    \ForAll {pairs $x,y \in X$}
    \State {Place $x, y$ into corresponding bucket }
    \EndFor
\EndFunction
\end{algorithmic}

We can use essentially the same proof as in \cite{bose2010computing}.
Claim 1. If $s > \lambda^2 $ and $A, B$ are $s$-separated, then all pairs $(a, a')$, $(b, b')$ with $a, a' \in A$
and $b, b' \in B$ are processed before any of the pairs $(a, b)$.

Indeed, if $s > \lambda^2$, then for every points $a_1, a_2 \in A$ and $b \in B$
we have $\lambda^2 d(a_1, a_2) < d(a_1, b)$. If $\dappr(a_1, a_2)$ and $\dappr(a_1, b)$ denote the $\lambda$-approximations
of the corresponding distances, then the inequality $\frac{\dappr(a_1, b)}{\dappr(a_1, a_2)} > \lambda$
holds , hence $(a_1, a_2)$ will be placed in the bucket that precedes the bucket of $(a_1, b)$.

Claim 2. If $s > \max(\lambda^2, \frac{2t+2}{t-1})$ and $A, B$ are two $s$-well-separated subsets of $X$,
then in the approximate greedy spanner computed by \ref{alg:approx_greedy_spanner} there is at most one edge between $A$ and $B$.

Suppose that the pair $(a_1, b_1)$ with $a_1 \in A$, $b_1 \in B$ was processed and the edge $(a, b)$
was added to the spanner. Then from claim 1 we know that we already have a $t$-spanner on $A$ and $B$.
Consider another pair $(a_2, b_2)$. Let $G$ be the weighted graph we have immediately after the edge $(a_1, b_1)$ has been added
to the spanner. By claim 1
$d_G(a_2, a_1) \leq td(a_2, a_1)$ and $d_G(b_1, b_2) \leq t d(b_1, b_2)$.
By properties of WSPD, $d_G(a_1, b_1) = d(a_1, b_1) \leq (1 + \frac{2}{s})d(a_2, b_2)$
and $t d(a_2, a_1) \leq \frac{t}{s} d(a_2, b_2)$, $t d(b_1, b_2) \leq \frac{t}{s} d(a_2, b_2)$.

Combining these inequalities, we get

\begin{equation}
    \begin{split}
        d_G(a_2, b_2) & \leq  d_G(a_2, a_1) + d_G(a_1, b_1) + d_G(b_1, b_2)  \\
                      & \leq  \frac{t}{s} d(a_2, b_2) + (1 + \frac{2}{s})d(a_2, b_2) + \frac{t}{s} d(a_2, b_2) \\
                      & =     (1 + \frac{2t + 2}{s}) d(a_2, b_2)
    \end{split}
\end{equation}

Solving for $s$ the inequality $1 + \frac{2t + 2}{s} < t$ gives $s > \frac{2t+2}{t-1}$.
